---
title: "A2_Writeup"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Call Option

According to the simulation, the price of a four upswing option is \$1.0625/liter of gasoline. Mr. Hamilton would have to pay \$53.125 for this price protection plan for 52 weeks period. 

As the graph below shows, the optimal exercising nodes are 
```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(rlist)
library(tidyverse)
library(reshape2)
library(ggplot2)
## Tree 1, underlying of oil price
iteration <- 1:53
underlying <- numeric(length = length(iteration))
track <- 1
expo <- 52
while (track <= 53){
  underlying[track] <- (1.1)^expo
  expo = expo - 2 
  track = track + 1
}


prev_value <- function(value, period){
  nweek = 1:period
  current_list = value 
  newlist = c()
  for (i in nweek){
    # newlist[i] <- ifelse(current_list[i] >= 1, current_list[i]/1.1, current_list[i]*1.1)
    newlist[i] <- current_list[i]/1.1
    
  }
  return(newlist)
}


recomb_tree <- function(value, n_period){
  output <- list(value)
  newlist <- prev_value(value, n_period)
  while(n_period >0) {
    output = list.append(output, newlist)
    n_period <- n_period -1 
    newlist <- prev_value(newlist, n_period)
  }
  return(output)
}

underlying_tree <- recomb_tree(underlying, 52)

# Find the intrinsic value of each node
intrinsic_value <- function (tree){
  output <- list()
  for (i in 1:length(tree)){
    newlist <- c()
    for (j in 1:length(tree[[i]])){
      newlist[j] <- ifelse(tree[[i]][j] -1 <0, 0, tree[[i]][j] -1) 
    }
    output = list.append(output, newlist)
  }
  return(output)
}

#intrinsic value of the underlying tree 
iv_tree <- intrinsic_value(underlying_tree)
```


```{r, echo = FALSE}
# Tree 2, value of American option (1 upswing)
## first calculate the European payout price 
payment <- numeric(length = length(iteration))
for (i in seq_along(iteration)){
  # After exercise option, how much is left over
  payment[i] = ifelse(underlying[i] -1 < 0, 0, underlying[i] -1)
}

## backward construction of 1 up swing call 
library(rlist)

backward <- function (value, nperiod){
  nweek <- 1:nperiod 
  scenario <-  matrix(c(1.1, 0.1, 0.909, -0.0909), nrow = 2, byrow = T)
  current_list <- value # store a list of option payout value
  newlist <- c()
  for (i in nweek){
    answer_i <- solve(scenario) %*% c(value[i], value[i + 1])
    newlist[i] <- answer_i[1]
  }
  newlist
}

tree_build <- function(t_value, nperiod){
  output <- list(t_value)
  newlist <- backward(t_value, nperiod)
  nperiod <- nperiod - 1 
  output <- list.append(output, newlist)
  # output <- list(newlist)
  while(nperiod > 0){
    newlist <- backward(newlist, nperiod)
    nperiod <- nperiod - 1
    output = list.append(output, newlist)
  }
  return(output)
}

## 1 up swing tree: 
upswing_1 <- tree_build(payment, 52)

```

 
```{r, echo= FALSE}
# 2- upswing
m_tree <- function(intrinsic ,old_tree, level = 53){
  payment2 <- old_tree[1]
  level = level - 1 
  start <- old_tree[[2]] + intrinsic[[2]]
  level = level - 1 
  new<- tree_build(start, level)
  for (i in 1:length(new)){
    payment2 <- list.append(payment2, new[[i]])
  }
  return(payment2)
}

two_upswing <- m_tree(iv_tree, upswing_1)

```


```{r, echo= FALSE}
# 3- upswing
m_tree3 <- function(intrinsic ,old_tree, level = 53){
  payment2 <- list(old_tree[[1]],old_tree[[2]])
  level = level - 2 
  start <- old_tree[[3]] + intrinsic[[3]]
  level = level - 1 
  new<- tree_build(start, level)
  for (i in 1:length(new)){
    payment2 <- list.append(payment2, new[[i]])
  }
  return(payment2)
}

three_up <- m_tree3(iv_tree, two_upswing)

```


 
```{r, echo= FALSE}
# 4- upswing
m_tree4 <- function(intrinsic ,old_tree, level = 53){
  payment2 <- list(old_tree[[1]],old_tree[[2]], old_tree[[3]])
  level = level - 3
  start <- old_tree[[4]] + intrinsic[[4]]
  level = level - 1 
  new<- tree_build(start, level)
  for (i in 1:length(new)){
    payment2 <- list.append(payment2, new[[i]])
  }
  return(payment2)
}

four_up <- m_tree4(iv_tree, three_up)
four_up[[53]]
```


```{r, echo= FALSE}
# Calculate Optimal Exercise Periods
opt_exercise_call <- function(tree_n, tree_n1, payment){
  
  exercise_nodes <- list()
  for (i in seq_along(tree_n)){
    new_list <- numeric(length = length(tree_n[[i]]))
    for (j in seq_along(tree_n[[i]])){
      if (payment[[i]][j] + tree_n[[i]][j] > tree_n1[[i]][j] & tree_n1[[i]][j] != 0 ){
        new_list[j] <- "exercise"
      } else if (payment[[i]][j] + tree_n[[i]][j] == tree_n1[[i]][j] & tree_n1[[i]][j] != 0){
        new_list[j] <- "indifferent"
      } else if ( payment[[i]][j] + tree_n[[i]][j] < tree_n1[[i]][j] & tree_n1[[i]][j] != 0 ) {
        new_list[j] <- "not_optimal"
      } else{
        new_list[j] <- "not_optimal"
      }
    }
    exercise_nodes = list.append(exercise_nodes, new_list)
  }
  return(exercise_nodes)
}

callnodes <- opt_exercise_call(three_up, four_up, iv_tree)
```

```{r, echo= FALSE}
# plot 
plot_machine <- function(nodes){
  storage <- c(rep(0, 53))
  for (i in seq_along(nodes)){
    vec <- nodes[[i]]
    if (length(vec) != 53){
        vec[length(vec):53] = "not_optimal"
    }
    storage <- cbind(storage, vec)
  }
  return(as.matrix(storage[,-1]))
}


nodescall <- plot_machine(callnodes)
colnames(nodescall) <- c(52:0)


ggplot(melt(nodescall), aes(y = Var1,x = Var2, fill= value)) +
  geom_tile() + 
  scale_fill_brewer(palette =  "Blues", direction =  -1)+
  labs(title = "Optimal Exercise Nodes in 4 up-swings", x = "time", y = "number of downs") + 
  scale_x_continuous(breaks = seq(0,52, 4))

```


\newpage
# Put Option

For the 4 downswing option, according to our simulation, it cost about \$1.0635/liter for the price protection plan. Thus, the price protection program would cost Ms. Curie \$53,175 to have the right to sell her gasoline at $50,000 for the 52 weeks period. 

```{r, echo = FALSE}
## Tree 1, underlying of oil price
iteration <- 1:53
underlying <- numeric(length = length(iteration))
track <- 1
expo <- 52
while (track <= 53){
  underlying[track] <- (1.1)^expo
  expo = expo - 2 
  track = track + 1
}


prev_value <- function(value, period){
  nweek = 1:period
  current_list = value 
  newlist = c()
  for (i in nweek){
    newlist[i] <- current_list[i]/1.1
  }
  return(newlist)
}


recomb_tree <- function(value, n_period){
  output <- list(value)
  newlist <- prev_value(value, n_period)
  while(n_period >0) {
    output = list.append(output, newlist)
    n_period <- n_period -1 
    newlist <- prev_value(newlist, n_period)
  }
  return(output)
}

underlying_tree <- recomb_tree(underlying, 52)

# Find the intrinsic value of each node
intrinsic_value_p <- function (tree){
  output <- list()
  for (i in 1:length(tree)){
    newlist <- c()
    for (j in 1:length(tree[[i]])){
      newlist[j] <- ifelse(1 - tree[[i]][j]  <0, 0, 1 - tree[[i]][j]) 
    }
    output = list.append(output, newlist)
  }
  return(output)
}

#intrinsic value of the underlying tree 
ivp_tree <- intrinsic_value_p(underlying_tree)
```



```{r, echo=FALSE}
# Tree2 American Option

backward_p <- function (value, nperiod){
  nweek <- 2:nperiod 
  scenario <-  matrix(c(1.1, 0.1, 0.909, -0.0909), nrow = 2, byrow = T)
  current_list <- value # store a list of option payout value
  newlist <- c()
  for (i in nweek){
    answer_i <- solve(scenario) %*% c(value[i], value[i + 1])
    newlist[i] <- answer_i[1]
  }
  newlist
}

tree_build_p <- function(t_value, nperiod){
  output <- list(t_value)
  newlist <- backward(t_value, nperiod)
  nperiod <- nperiod - 1 
  output <- list.append(output, newlist)
  # output <- list(newlist)
  while(nperiod > 0){
    newlist <- backward(newlist, nperiod)
    nperiod <- nperiod - 1
    output = list.append(output, newlist)
  }
  return(output)
}

downswing_1 <- tree_build_p(ivp_tree[[1]], 52)
downswing_1[[53]]
```


```{r, echo=FALSE}
# Tree3 2downswing 
twodown_tree <- function(intrinsic ,old_tree, level = 53){
  payment2 <- old_tree[1]
  level = level - 1 
  start <- old_tree[[2]] + intrinsic[[2]]
  level = level - 1 
  new<- tree_build(start, level)
  for (i in 1:length(new)){
    payment2 <- list.append(payment2, new[[i]])
  }
  return(payment2)
}

two_downswing <- m_tree(ivp_tree, downswing_1)

```




```{r, echo = FALSE}
# Tree4 3downswing
threedown_tree <- function(intrinsic ,old_tree, level = 53){
  payment2 <- list(old_tree[1], old_tree[2])
  level = level - 2 
  start <- old_tree[[3]] + intrinsic[[3]]
  level = level - 1 
  new<- tree_build(start, level)
  for (i in 1:length(new)){
    payment2 <- list.append(payment2, new[[i]])
  }
  return(payment2)
}

three_down <- threedown_tree(ivp_tree, two_downswing)

```


```{r, echo= FALSE}
# Tree5 4downswing
fourdown_tree <- function(intrinsic ,old_tree, level = 53){
  payment2 <- list(old_tree[[1]],old_tree[[2]], old_tree[[3]])
  level = level - 3
  start <- old_tree[[4]] + intrinsic[[4]]
  level = level - 1 
  new<- tree_build(start, level)
  for (i in 1:length(new)){
    payment2 <- list.append(payment2, new[[i]])
  }
  return(payment2)
}

four_down <- m_tree4(ivp_tree, three_down)
four_down[[53]]
```

 
```{r, echo= FALSE}
# Calculate put exercise period
opt_exercise_put <- function(tree_n, tree_n1, payment){
  
  exercise_nodes <- list()
  for (i in seq_along(tree_n)){
    new_list <- numeric(length = length(tree_n[[i]]))
    for (j in seq_along(tree_n[[i]])){
      if (payment[[i]][j] + tree_n[[i]][j] > tree_n1[[i]][j] & tree_n1[[i]][j] != 0 ){
        new_list[j] <- "exercise"
      } else if (payment[[i]][j] + tree_n[[i]][j] == tree_n1[[i]][j] & tree_n1[[i]][j] != 0){
        new_list[j] <- "indifferent"
      } else if ( payment[[i]][j] + tree_n[[i]][j] < tree_n1[[i]][j] & tree_n1[[i]][j] != 0 ) {
        new_list[j] <- "not optimal"
      } else {
        new_list[j] <- "not optimal"
      }
    }
    exercise_nodes = list.append(exercise_nodes, new_list)
  }
  return(exercise_nodes)
}

putnodes <- opt_exercise_put(three_down, four_down, ivp_tree)
```



```{r, echo= FALSE}
plot_machine_put <- function(nodes){
  storage <- c(rep(0, 53))
  for (i in seq_along(nodes)){
    time <- nodes[[i]]
    if (length(time) != 53){
      b = rep("not optimal", (53 - length(time)))
      time <- append(b, time)
    }
    storage <- cbind(storage, time)
  }
  return(as.matrix(storage[, -1]))
}

put_exercise <- plot_machine_put(putnodes)
colnames(put_exercise) <- c(52:0) # columns are time
ggplot(melt(put_exercise), aes(y = Var1,x = Var2, fill= value)) +
  geom_tile() + 
  scale_fill_brewer(palette = "Blues", direction =  -1) +
  labs(title = "Optimal Exercise Nodes in 4 down-swings", x = "time", y = "number of downs") + scale_x_continuous(breaks = seq(0,52, 4))
  

```





